Quantum Mechanics Assignment Eight

### 2016-03-01 quantum/latex/a8.tex
work on a8

@@

\begin{align*}
  \braket{(x_1-x_2)^2} &= \bra{n}_1\bra{k}_2(x_1-x_2)^2\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1\bra{k}_2(x_1^2 + x_2^2 - 2x_1x_2)\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1\bra{k}_2x_1^2\ket{n}_1\ket{k}_2 +
  \bra{n}_1\bra{k}_2x_2^2\ket{n}_1\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1x_1^2\ket{n}_1\bra{k}_2\ket{k}_2 +
  \bra{n}_1\ket{n}_1\bra{k}_2x_2^2\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1x_1^2\ket{n}_1 +
  \bra{k}_2x_2^2\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  now look up the first two terms and turn the third into an integral\\
\end{align*}

>>>>

\section{Problem 2: Neutron Star}
Neutron stars are celestial objects made up almost exclusively of neutrons, or spin-1/2 fermions.\\

@@

\begin{align*}
  \braket{(x_1-x_2)^2} &= \bra{n}_1\bra{k}_2(x_1-x_2)^2\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1\bra{k}_2(x_1^2 + x_2^2 - 2x_1x_2)\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1\bra{k}_2x_1^2\ket{n}_1\ket{k}_2 +
  \bra{n}_1\bra{k}_2x_2^2\ket{n}_1\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1x_1^2\ket{n}_1\bra{k}_2\ket{k}_2 +
  \bra{n}_1\ket{n}_1\bra{k}_2x_2^2\ket{k}_2 - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\bra{n}_1
  \left(a^\dagger a^\dagger + aa + aa^\dagger + a^\dagger a\right)\ket{n}_1
  + \frac{\hbar}{2m\omega}\bra{k}_2
  \left(a^\dagger a^\dagger + aa + aa^\dagger + a^\dagger a\right)\ket{k}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\bra{n}_1\left(aa^\dagger + a^\dagger a\right)\ket{n}_1
  + \frac{\hbar}{2m\omega}\bra{k}_2\left(aa^\dagger + a^\dagger a\right)\ket{k}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
  \left(a_1+a_1^\dagger\right)\left(a_2+a_2^\dagger\right)\ket{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
  \left(a_1a_2 + a_1^\dagger a_2^\dagger + a_1a_2^\dagger + a_2a_1^\dagger\right)\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)
  - \frac{2\hbar}{2m\omega}\bra{n}_1\bra{k}_2
  \left(a_1a_2 + a_1^\dagger a_2^\dagger + a_1a_2^\dagger + a_2a_1^\dagger\right)\bra{n}_1\ket{k}_2\\
  &= \frac{\hbar}{2m\omega}\left(2n + 2k + 2\right)\\
  &- \frac{2\hbar}{2m\omega}
  \left(\bra{n}_1a_1\ket{n}_1\ket{k}_2a_2\ket{k}_2
  + \bra{n}_1a_1^\dagger\ket{n}_1\bra{k}_2a_2^\dagger\ket{k}_2
  + \bra{n}_1a_1\ket{n}_1\bra{k}_2a_2^\dagger\ket{k}_2
  + \bra{k}_2a_2\ket{k}_2\bra{n}_1a_1^\dagger\ket{n}_1\right)\\
  &= \frac{\hbar}{m\omega}(2n+2k+2)\\
\end{align*}

\textbf{b.) Boson}

\begin{align*}
  \braket{(x_1-x_2)^2} &=
  \frac{1}{2}\left(\bra{n}_1\bra{k}_2 + \bra{k}_1\bra{n}_2\right)
  \left(x_1^2 + x_2^2 - 2x_1x_2\right)
  \left(\ket{n}_1\ket{k}_2 + \ket{k}_1\ket{n}_2\right)\\
  &= \frac{1}{2}\left(
  \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
  + \bra{k}_1x_1^2\ket{k}_1 + \bra{n}_2x_2^2\ket{n}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 - 2\bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\right)\\
  &= \frac{1}{2}\left(4n + 4k + 4\right)
  - \bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 - \bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\\
  &= \frac{\hbar}{2m\omega}\left(4n + 4k + 4\right)\\
\end{align*}

\textbf{c.) Fermion}

\begin{align*}
  \braket{(x_1-x_2)^2} &=
  \frac{1}{2}\left(\bra{n}_1\bra{k}_2 - \bra{k}_1\bra{n}_2\right)
  \left(x_1^2 + x_2^2 - 2x_1x_2\right)
  \left(\ket{n}_1\ket{k}_2 - \ket{k}_1\ket{n}_2\right)\\
  &= \frac{1}{2}\left(
  \bra{n}_1x_1^2\ket{n}_1 + \bra{k}_2x_2^2\ket{k}_2
  - \bra{k}_1x_1^2\ket{k}_1 - \bra{n}_2x_2^2\ket{n}_2
  - 2\bra{n}_1\bra{k}_2x_1x_2\ket{n}_1\ket{k}_2 + 2\bra{n}_2\bra{k}_1x_1x_2\ket{n}_2\ket{k}_1\right)\\
  &= \frac{\hbar}{2m\omega}\frac{1}{2}\left(2n+1 + 2k+1 - 2n-1 - 2k-1 - 0\right)\\
  &= 0\\
\end{align*}

This is a really interesting problem that demonstrates how systems that are identical save
for their symmetry rules can behave totally differently. The boson system has twice the expectation
value of the distinguishable particle system, and the fermion system has zero expectation value
for particle distance. This is a very unintuitive result, since it predicts that on average the
two particles are on top of each other.\\

\section{Problem 4: Three-particle system}
In this problem we practice writing the wave functions for three-particle systems of particles
with different symmetry rules:\\

\section{Problem 5: Zeeman Effect}
